---
title: "re_stepwise"
author: "Reto Moser"
date: "2023-10-04"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(ggplot2)
library(dplyr)
```

##Task

In this Task I built a stepwise forward regression and model the GPP as a function of predictors in the dataset of half-hourly ecosystem fluxes. GPP is the total amount of CO2 fixed by land plants per unit time (gross primary production). First I loaded the necessaty libraries and the data. Looking at the dataset shows which predictor variables are available. In a next step, we define the predictor and response variables. 

###Preparing data

```{r}

fluxes <- read_csv("data/df_for_stepwise_regression.csv")
spec(fluxes)


response <- fluxes$GPP_NT_VUT_REF
predictors <- c("TA_F", "SW_IN_F", "LW_IN_F", "VPD_F", "PA_F", "P_F", "WS_F", "TA_F_MDS", "SW_IN_F_MDS", "LW_IN_F_MDS", "VPD_F_MDS", "CO2_F_MDS", "PPFD_IN", "USTAR")
print(predictors)
```

##Evaluation of all bivariate models (single predictor)

To calculate the Regression with all the predictors only using one after another without connecting them together I used an empty vector and a for-loop to calculate the R-squared of each predictor/response variable. After the highest R-squared value is taken and the linear regression calculated again I got the AIC and the name of the variable. In this case it is the variable for the photosynthetic photon flux density (PPFD_IN).

```{r}
rsquare <- c()
for (predictor in predictors) {
  formel <- formula(paste("GPP_NT_VUT_REF ~", predictor))
  model <- lm(formel, data = fluxes)
  summary_model <- summary(model)
  print(summary_model)
  r_square <- summary_model$r.squared
  rsquare <- c(rsquare, r_square)
  print(rsquare)
}




highest_r2 <- which.max(rsquare)
name <- predictors[highest_r2]
formel_2 <- formula(paste("GPP_NT_VUT_REF ~", name))
model_2 <- lm(formel_2, data = fluxes)
aic <- AIC(model_2)
print(rsquare)
cat("Lowest AIC:", aic, 
    "Variable:", name)
```

#increase to p + 1

```{r}
response <- fluxes$GPP_NT_VUT_REF
predictors <- c("TA_F", "SW_IN_F", "LW_IN_F", "VPD_F", "PA_F", "P_F", "WS_F", "TA_F_MDS", "SW_IN_F_MDS", "LW_IN_F_MDS", "VPD_F_MDS", "CO2_F_MDS", "PPFD_IN", "USTAR")

small_aic <- Inf
model_3 <- NULL

for (i in 1:14) {
     combinations <- combn(predictors, i)
     for (j in 1:ncol(combinations)){
       formel_p <- formula(paste("GPP_NT_VUT_REF ~", paste(combinations[,j], collapse = "+")))
       model_3lm <- lm(formel_p, data = fluxes)
       aic <- AIC(model_3lm)
       if (aic < small_aic){
         small_aic <- aic
         model_3 <- model_3lm
       } else {
         break
     }
    }
  }

summary(model_3lm)
print(small_aic)

```


