---
title: "re_tidy"
author: "Reto Moser"
date: "2023-03-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Loading Data, defining variables and group data

I load the necessary packages for this report exercise.

In the .csv file I cleaned all the unnecessary data by hand.

With the read_csv2() command and semicolons as delimiter I loaded the .csv document. After reviewing it, the variable "time" had to be changed from character to numeric for the following tasks. With the command is.numeric I check if the change worked and if the variable time is now numeric.

Then I grouped the data by experiment.

```{r read data}

library(tidyverse)
library(readr)
library(dplyr)

data_S1 <- read_csv2("data/AGDS_report_1_tidy.csv")
data_S1_head <- head(data_S1, n = 3)
data_S1_head

data_S1 <- data_S1 %>% mutate(time = as.numeric(time))
is.numeric(data_S1$time)


group_S1 <- group_by(data_S1, experiment)
group_S1

```

## Calculate log-response-ratio 

Instead of calculating the log response for each row in the .csv file and adding it to the file using the mutate() command, we need to obtain a single log-response ratio for each experiment. This means we should have a total of only 53 log-response ratios. To achieve this, we can use the summarize() command to summarize the rows for each experiment and calculate the mean log-response ratio, which is the desired outcome.

```{r}
group_S1_RR <- mutate(group_S1, RR = log(co2_increased_mean / co2_ambient_mean))
group_S1_RR

experiment_RR <- summarize(group_S1_RR, mean_log_ratio = mean(log(co2_increased_mean / co2_ambient_mean)))
knitr::kable(experiment_RR)
```

## Aggregtae data into three time phases

By creating a column named "Phase" I can group the data by years. I did this with the cut() command breaking the time-column at 0,3,6, and Infinity. Then labeling them as Early, Mid and Late phase. 

- early phase (<3 years since start)
- middle phase (3-6 years since start)
- late phase (>6 years since start)

Then I calculated the log-response-ratio for the three phases and then I aggregated the log-respond-ratios by taking their mean. 

```{r}
data_S1_RR <- mutate(data_S1, RR = log(co2_increased_mean / co2_ambient_mean))
data_S1_RR

data_S1_RR$Phase <- cut(data_S1_RR$time, breaks = c(0, 3, 6, Inf), 
                       labels = c("Early phase", "Mid phase", "Late phase"), 
                       include.lowest = TRUE)
data_S1_RR
data_S1_RR_head <- head(data_S1_RR, n = 3)
knitr::kable(data_S1_RR_head)


mean_RR_by_phase <- tapply(data_S1_RR$RR, data_S1_RR$Phase, mean)
mean_RR_by_phase
```




